---
title: "NormalyzerDE analysis of Bile Acid Proteomics Data"
output:
  pdf_document: 
    highlight: tango
  html_notebook: default
---
 
```{r lib-1, echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(limma)
library(edgeR)
library(readxl)
```

# Maxquant Object Import

Read in the MaxQuant proteinGroups file. 
Only keep columns with iBAQ intensitites, gene info, and contaminant info.
Remove iBAQ peptides column from the protein groups file.
```{r eval=FALSE}
mq <- read_excel("ravi_pg.xlsx", col_types = c("text", 
     "text", "text", "text", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "numeric", "numeric", "numeric", 
     "numeric", "text", "text", "text"))
mq_cleaned <- mq %>% 
  filter(is.na(Reverse) , is.na(`Only identified by site`), is.na(`Potential contaminant`)) %>% 
  select(`Protein IDs`, `Majority protein IDs`, `Protein names`, `Gene names` ,matches("iBAQ (([01][0-9])|(2[0-4]))")) 
codes <- read_excel(here::here("ravi_proteinGroups.xlsx"), sheet = 2) %>% 
  mutate(reps = case_when(
    str_detect(identity, "C.c") ~ "Cc",
    str_detect(identity, "C.m") ~ "Cm",
    str_detect(identity, "L.c") ~ "Lc",
    str_detect(identity, "L.m") ~ "Lm"
  ))
```

Prepare tables for normalyzer analysis.

```{r eval=FALSE}
#Prepare objects for normalyzer
norm_meta <- codes %>% 
  select("sample" = identity, "group" = reps)
norm_mq <- mq_cleaned %>% 
  select(`Majority protein IDs`, contains("iBAQ")) %>% 
  rename_with(function(x) norm_meta$sample [1:24], contains("iBAQ")) %>% 
  filter(!str_detect(`Majority protein IDs`,"REV_"))
write_tsv(norm_meta [1:24,], file = "norm_meta.tsv")
write_tsv(norm_mq , file = "norm_mq.tsv")
```

# NormalyzerDE: normalization

Use normalyzerDE to test all normalization methods.
 
```{r lib-2, echo=TRUE, warning=FALSE, message=FALSE}
library(NormalyzerDE)
library(SummarizedExperiment)
```

```{r eval=FALSE}
normalyzer(jobName="test_run", designPath="norm_meta.tsv", dataPath="norm_mq.tsv",
           outputDir = here::here("normalyzer/"))

```

# Differential Abundance Analysis

Based on the QC plots from normalyzer DE, use vsn normalization for 
differential abundance analysis.

```{r eval=FALSE}
normalyzerDE("test_run", designPath = here::here("norm_meta.tsv"), 
             dataPath=here::here("normalyzer/test_run/VSN-normalized.txt"),
             outputDir = here::here("normalyzer/"), comparisons = c("Cc-Lc","Cm-Lm"),
             sigThres = 0.05, log2FoldThres = 1)
```

Combine the original intensities table, and the DA analysis output.

```{r eval=FALSE}
norm_meta <- read_tsv(here::here("norm_meta.tsv"))
norm_mq_annot <- mq_cleaned %>% 
  select(-contains("iBAQ")) 
res <- read_tsv(here::here("normalyzer/test_run/test_run_stats.tsv"))
#Annotate the result table
res_annot <- res %>% 
  left_join(norm_mq_annot)
write_tsv(res_annot, file = here::here("normalyzer_limma_de_results_annotated.tsv"))
```

# Session Info

```{r}
sessionInfo()
```


